\documentclass{article}

\begin{document}

\begingroup
  \centering
  \textbf{\large Welcome to the \confYear\ CCSC \confName\ Conference\\}
\endgroup

\vspace{10pt}

The 2023 South Central Steering Committee is very pleased to welcome everyone to our 34th annual conference hosted by Stephen F. Austin State
University in Nacogdoches, Texas. Unlike the previous two years, our conference chair and host, Anne Marie Eubanks, has provided infrastructure and support
for the in-person delivery of our conference this year.

For our 2023 conference, we have seven papers, three tutorials, and both student and faculty posters scheduled for the program. This year the Steering
Committee chose 7 of 15 papers through a double-blind review process for a paper acceptance rate of 46\%. Sixteen colleagues across the region and country
served as professional reviewers and we recognize the expertise and guidance they all so thoughtfully contributed to the selection of our 2023 conference
program.

The Steering Committee continues to seek colleagues to host the conference in the future and to join our community of computer science educators to enrich
our curricula and provide innovative pedagogy for our students. We invite and encourage our fellow members of the South Central region to attend our steering
committee business meeting on Friday, March 31, 2023 after the conference reception and banquet. Fellow educators and colleagues are encouraged to join
in our efforts to involve more of our community in the planning and execution of the conference in the future.

We extend a very warm and delightful welcome to our presenters and attendees who continue to promote computer science education and camaraderie to
our region. To all members of our 2023 Steering Committee, thank you again for your help in organizing the conference and your gracious efforts in delivering our conference during such challenging times.

\vspace{10pt}

\hfill Anne Marie Eubanks

\hfill Stephen F. Austin State University

\hfill Conference Chair and Host

\vspace{10pt}

\hfill Bingyang Wei

\hfill Texas Christian University

\hfill Regional Editor Co-Chair

\vspace{10pt}

\hfill Mustafa Al-Lail

\hfill Texas A\&M International University

\hfill Regional Editor Co-Chair
\end{document}
